---
title: "Fiscal policy shocks and stock prices"
author: "Hanwen Zhang"

execute:
  echo: false
  
bibliography: references.bib
---

> **Abstract.** This is a reserarch proposal, the research is about measuring fiscal policy effect using Structural VAR models in Australian Economy.
>
> **Keywords.** fiscal policy shock, SVAR, tax shocks, stock price

# Introduction

How does tax cuts affect the economy? Will tax reductions boost stock market? This report aims to answer these questions in the background of Australian Economy. These questions are crucial to understanding of dynamic economics and the stimulus policy-making process. 

# Literature Review
Estimated effects of tax shocks on output don't agree on each other in the previous literature. 


# Data 
The variables that are of interest include GDP, total government expenditure, total government revenue, interest rate, inflation, public debt and stock prices. **Justify.** 

```{r collecting data of interest}
#| echo: false
#| message: false
#| warning: false

# quarterly, seasonally adjusted, nominal, GDP per capita, unit: $
# 1973-09-01 to 2022-12-01 --> 198 quarters 
gdp_no_pc <- readabs::read_abs(series_id = "A2304420C") 
tmp.gdp_no_pc <- na.omit(xts::xts(gdp_no_pc$value, gdp_no_pc$date) )
# head(na.omit(tmp.gdp_no_pc)) 

# quarterly, seasonally adjusted, nominal, GDP, unit: $M
# 1959-09-01 to 2022-12-01 --> 254 quarters 
gdp_no <- readabs::read_abs(series_id = "A2304418T") 
tmp.gdp_no <- xts::xts(gdp_no$value, gdp_no$date) 
# length(na.omit(tmp.gdp_no)) 

# Get population unit: M
pop <-  tmp.gdp_no/tmp.gdp_no_pc

# quarterly, seasonally adjusted, GDP deflator, unit: 
# 1959-09-01 to 2022-12-01 --> 254 quarters 
gdp_deflator <- readabs::read_abs(series_id = "A2303730T") 
tmp.gdp.deflator <- xts::xts(gdp_deflator$value, gdp_deflator$date) 

# construct log real GDP per capita
gdp_real_pc <- na.omit(log(tmp.gdp_no_pc/tmp.gdp.deflator)) # 1973-09-01 to 2022-12-01

# quarterly, seasonally adjusted, government spending, unit: $M
# 1959-09-01 to 2022-12-01 --> 254 quarters 
gov_spend <- readabs::read_abs(series_id = "A2304036K") 
tmp.gov_spend  <- xts::xts(gov_spend$value, gov_spend$date) 

# quarterly, original, government spending, unit: $M
# 1959-09-01 to 2022-12-01 --> 254 quarters 
gov_revenue <- readabs::read_abs(series_id = "A2302411R") 
tmp.gov_revenue <- xts::xts(gov_revenue$value, gov_revenue$date) 


# original CPI
# 1948-09-01 to 2022-12-01 --> 298 quarters 
cpi <-  readabs::read_abs(series_id = "A2325846C") 
tmp.cpi <- xts::xts(cpi$value, cpi$date) 
# length(tmp.cpi)
# tail(tmp.cpi)

inflation <- 100*diff(tmp.cpi)/tmp.cpi

# readrba::read_rba(series_id = "FIRMMCRI") 
# A2149474A


# quarterly, seasonally adjusted, Total taxes, unit: $M
# 1959-09-01 to 2022-12-01 --> 254 quarters 
# from table: 5206022_Taxes A2302794K

```


```{r data visualization}
plot.ts(gdp_real_pc,main="Log Quarterly Real GDP Per Capita")
```


# Methodology

Structural vector autoregression (SVAR) models can be used to estimate the impact of taxation shocks in Australian economy. The benchmark model is defined as: 
\begin{align}
B_0Y_t = b_0 + \sum_{i=1}^{p}B_{i}Y_{t-i}+u_t 
\end{align}
\begin{align}
u_t|Y_{t-1}\sim iid(0_N,I_N)
\end{align}

$Y_t$ is $N\times1$ matrix of endogeneius variables, N represents the number of endogeneous variables and p is the number of lags, $B_0$ is $N\times N$ matrix capturing contemporaneous relationships between variables, and $u_t$ is $N\times1$ vector conditionally on $Y_{t-1}$ orthogonal structural shocks. 

The reduced form representation is: 
\begin{align}
Y_t = \mu_0 + \sum_{i=1}^{p}A_{i}Y_{t-i}+\epsilon_t
\end{align}
\begin{align}
\epsilon_t|Y_{t-1}\sim iid(0_N,\Sigma)
\end{align}

The covariance matrix of $\epsilon_t$ can be written as:
\begin{align}
\Sigma=B_0^{-1}{B_0^{-1}}^{'}
\end{align}

# Preliminary Results

# References {.unnumbered}
